Kernel Phase and Coronagraphy with Automatic Differentiation

### 2020-05-19 ms.tex
Updates from Overleaf

Arbitrary-order Taylor series expansions for a PSF with respect to small phase perturbations can be derived analytically for a telescope with an arbitrary pupil brought to a single focus \citep{bloemhof01,anand02,perrin03}. This expansion breaks down for more complex optical systems. The PSF of a coronagraphic imager far from the occulted region is very similar to that of the primary PSF of the telescope, but close to the inner working angle it is significantly affected by the occultation and especially strongly by low-order wavefront errors. This is likewise an issue where detector nonlinearity and saturation significantly distort the PSF While the analytic series expansion is no longer applicable in these regimes, using matrix Fourier transforms and skipping saturated pixels or pixels inside the coronagraphic inner working angle has shown promise for pushing kernel phase beyond its conventional limitations \citep{laugier19b}.

@@

\subsection{Simple Diffraction}
\ref{sec:simple}

@@

\begin{figure}
    \plotone{kerphi_singular.pdf}
     \caption{Log-scale plot of the singular values of the phase transfer matrix versus index for simple monochromatic diffraction (blue), 20\% bandwidth (green), and a monochromatic Lyot coronagraph (orange). Singular values are normalized to the first singular value. In both simple diffraction cases, there is a sharp cutoff separating significant singular values from a subset within machine precision of zero. For the Lyot coronagraph the roll-off is much smoother, but nevertheless reaches small singular values. The vectors corresponding to these approximately zero singular values span the null space of the phase transfer operator, and for the simple diffraction case and coronagraphic case respectively create traditional kernel phases and kernel speckles. The coronagraphic case  exhibits a slow roll-off for a much larger region than in the case of the full pupil. }
    \label{fig:svd}
\end{figure}

@@

We have also repeated the above calculations for light with a $20\%$ fractional bandwidth\footnote{\href{https://github.com/benjaminpope/morphine/blob/master/notebooks/morphine_u,v_broadband.ipynb}{github.com/benjaminpope/morphine/blob/master/notebooks/morphine\_u,v\_broadband.ipynb}}, with a uniform spectrum sampled ten times from $1.8\,\mu\text{m}$\,---\,$2.2\,\mu\text{m}$. The results are very similar, except that the Jacobian maps equivalent to Figure~\ref{kernel_jacobian} are slightly blurred. The number of singular values (shown in green in Figure~\ref{fig:svd}) differs from the monochromatic case because the $u,v$ support is slightly changed in the broadband case, but overall a qualitatively similar behaviour is found between the broadband and monochromatic cases.

@@

We first build a high resolution model of the PHARO pupil using the same code as \citet{martinache20}, and bin it down to a $64\times64$ pixel grid so that pixel values $\in [0,1]$ indicate the fraction of the telescope aperture filling that binned pixel. We then propagate this onto a $64\times64$-pixel image plane, and use a matrix Fourier transform to map this onto a $u,v$ plane of the same size. We differentiate this using the forwards mode autodiff as above, obtaining the Jacobian shown in Figure~\ref{pharo_jacobian}, and calculate a kernel phase transfer matrix.

@@

% other autodiff packages - TensorFlow was slower?
Because of its \textsc{NumPy}-like API, we have used \textsc{jax} to power the automatic differentiation in \textsc{morphine}. We have not extensively optimized this code for the most natural autodiff, nor have we benchmarked it against competing frameworks such as \textsC{TensorFlow}. We expect that integration with \textsc{NumPy} and similarity to the existing standard \textsc{poppy} will make \textsc{morphine} more readily useful in astronomy, but for integration with neural networks it may be more convenient to use a different autodiff package.

@@

In the interests of open science, we have made available the Python package \textsc{morphine}, together with Jupyter notebooks used to generate the figures in this paper, under a BSD 3-clause open source license at \href{https://github.com/benjaminpope/morphine}{github.com/benjaminpope/morphine}.

@@

%outside of astronomy
The generalization of kernel phase and optical gradient design permitted by autodiff also allows us to extend this work to the near-field (Fresnel) propagation regime. This may lead to improvements in modelling of components (such as deformable mirrors) that might occur in planes intermediate between pupil and focus. This may be most relevant outside of astronomy: for example, Fresnel coherent diffractive imaging \citep{williams2006} is a popular microscopy technique, in which gradient-based advances in phase retrieval, kernel phase and optimization may be valuable.

>>>>

Arbitrary-order Taylor series expansions for a PSF with respect to small phase perturbations can be derived analytically for a telescope with an arbitrary pupil brought to a single focus \citep{bloemhof01,anand02,perrin03}. This expansion breaks down for more complex optical systems. The PSF of a coronagraphic imager far from the occulted region is very similar to that of the primary PSF of the telescope, but close to the inner working angle it is significantly affected by the occultation and especially strongly by low-order wavefront errors. This is likewise an issue where detector nonlinearity and saturation significantly distort the PSF. While the analytic series expansion is no longer applicable in these regimes, using matrix Fourier transforms and skipping saturated pixels or pixels inside the coronagraphic inner working angle has shown promise for pushing kernel phase beyond its conventional limitations \citep{laugier19b}, as have generalizations that exploit angular differential imaging for a further level of calibration \citep{laugier20}.

@@

\subsection{Simple Diffraction}
\label{sec:simple}

@@

\begin{figure}
    \plotone{kerphi_singular.pdf}
     \caption{Log-scale plot of the singular values of the phase transfer matrix versus index for simple monochromatic diffraction (blue), 20\% bandwidth (green). Singular values are normalized to the first singular value. In both cases, there is a sharp cutoff separating significant singular values from a subset within machine precision of zero. The vectors corresponding to these approximately zero singular values span the null space of the phase transfer operator, and correspond to kernel phases.}
    \label{fig:svd}
\end{figure}

@@

We have also repeated the above calculations for light with a $20\%$ fractional bandwidth\footnote{\href{https://github.com/benjaminpope/morphine/blob/master/notebooks/morphine_u,v_broadband.ipynb}{github.com/benjaminpope/morphine/blob/master/notebooks/morphine\_u,v\_broadband.ipynb}}, with a uniform spectrum sampled ten times from $1.8\,\mu\text{m}$\,---\,$2.2\,\mu\text{m}$. The results are very similar, except that the Jacobian maps equivalent to Figure~\ref{kernel_jacobian} are slightly blurred. In the shape of the singular value curve between the broadband and monochromatic cases, a qualitatively similar behaviour is found, although the broadband case has slightly fewer kernel phases and a flatter roll-off.

@@

We first build a high resolution model of the PHARO pupil using the same code as \citet{martinache20}, and bin it down to a $64\times64$ pixel grid so that pixel values $\in [0,1]$ indicate the fraction of the telescope aperture filling that binned pixel. We then propagate monochromatic 2.145\,$\mu$m light through this onto a $64\times64$-pixel image plane, and use a matrix Fourier transform to map this onto a $u,v$ plane of the same size. We differentiate this using the forwards mode autodiff as above, obtaining the Jacobian shown in Figure~\ref{pharo_jacobian}, and calculate a kernel phase transfer matrix.

@@

It will be valuable to consider whether the small difference between the $\alpha$~Ophiuchi binary parameters between this and previous work is due to a differing sensitivity to systematics in one or the other pipeline, or whether due to a flaw in the way the kernel phase matrix is calculated in the autodiff case. Fourier sampling, and numerical noise in the Jacobian, both present the possibility of suboptimal performance, though the method appears to perform well on simulated data.

% other autodiff packages - TensorFlow was slower?
Because of its \textsc{NumPy}-like API, we have used \textsc{jax} to power the automatic differentiation in \textsc{morphine}. We have not extensively optimized this code for fast and memory-efficient autodiff, nor have we benchmarked it against competing frameworks such as \textsC{TensorFlow}. We expect that integration with \textsc{NumPy} and similarity to the existing standard \textsc{poppy} will make \textsc{morphine} more readily useful in astronomy, but for integration with neural networks it may be more convenient to use a different autodiff package.

@@

In the interests of open science, we have made available the Python package \textsc{morphine}, together with Jupyter notebooks used to generate the figures in this paper, under a BSD 3-clause open source license at \href{https://github.com/benjaminpope/morphine}{github.com/benjaminpope/morphine}. We encourage and welcome other scientists to replicate, apply, and extend our work.

@@

%outside of astronomy
The generalization of kernel phase and optical gradient design permitted by autodiff also allows us to extend this work to the near-field (Fresnel) propagation regime. This may lead to improvements in modelling of components (such as deformable mirrors) that might occur in planes intermediate between pupil and focus, or simply more accurate kernel phase models of instruments such as the \textit{Hubble Space Telescope}. This may be most relevant outside of astronomy: for example, Fresnel coherent diffractive imaging \citep{williams2006} is a popular microscopy technique, in which gradient-based advances in phase retrieval have been applied \citep{Dueaay3700}, and kernel phase and optimization may be valuable.
